### 2024-01-17 nondirectional_ms_2024.Rmd
Added data availabiltiy note + changed appendices order + replaced SM with appendix

authornote: |
  All data, and analysis code are available at [https://github.com/mufcItay/NDT]. This study's analysis was not pre-registered.

@@

# the analysis and simulations for appendix B.
# to run the analysis again, uncomment the (commented) run_appendixB line and comment the read.csv line which reads the results of an existing simulations run.
source('appendix\\appendix_B.R')
# appendixB_res <- run_appendixB(apndxB_conf_FAs, apndxB_conf_sensitivity)
appendixB_res <- read.csv('appendix\\results\\2024-01-07_23_39_16Appendix_B_QUID_OANOVA.csv')

# the analysis and simulations for appendix C.
# to run the analysis again, uncomment the (commented) run_appendixC line and comment the read.csv line which reads the results of an existing simulations run (keeping the save_plot_appendixC line.)
source('appendix\\appendix_C.R')
# appendixC_res <- run_appendixC(apndx_b_conf)
appendixC_res <- read.csv('appendix\\results\\2023-12-29_18_48_11Appendix_C.csv')
save_plot_appendixC(appendixC_res)

# appendix D was not implemented in an external script.
# see the implementation below

# the materials for appendix E are generated by sourcing the
# 'figure4.R' script, and the its figure was created with minor
# edits applied to the output files

# the analysis and simulations for appendix F.
# to run the analysis again, uncomment the (commented) run_appendixF line and comment the read.csv line which reads the results of an existing simulations run.
source('appendix\\appendix_F.R')
apndxF_N_p <- N_p
apndxF_N_t <- N_t
# appendixF_res <- run_appendixF(apndx_e_conf)
appendixF_res <- read.csv('appendix\\results\\2024-01-09_16_46_09Appendix_F.csv')
apndx_F_median_cnd_power <- appendixF_res %>%
  filter(N_p == median(appendixF_res$N_p), 
         N_t == median(appendixF_res$N_t),
         sds_ratio == median(appendixF_res$sds_ratio)) %>%
  pull(Power)
```

@@

We had three inclusion criteria: first, since all probed tests require trial-level data, only open-access datasets providing such data were included. Second, the independent variable of all studies had to be manipulated within single participants. And third, at least one non-significant effect was reported in the original study. Overall, this search strategy yielded data associated with `r n_empiric_ns` null effects (see Supplementary Table 1 for details about all effects), 21 focusing on differences in RT and 5 on differences in signal detection sensitivity, d’ [@green1966a]. We used the criteria set by the original authors for demonstrating unawareness (e.g., using objective and/or subjective measures of awareness), and a two-sided non-parametric sign-flipping test on the population mean for filtering out significant priming effects^[Across all RT effects, our analysis used raw RT scores, and thus our results diverged from the original results when log transformations were used (see the notes column in Supplementary Table 1 for details)]. Finally, we excluded participants with fewer than five trials per experimental condition and/or zero variance in the dependent variable (e.g., when accuracy was measured). Together, these data allowed us to reexamine null unconscious processing effects using a non-directional approach that takes into account the potential for differences in effect signs when testing for group-level effects. We accordingly asked whether true effects of unconscious processing were masked by population heterogeneity in effect signs.

To that end, the effects of interest were tested using GNT, QUID, and the OANOVA tests (see Supplementary Figure 1 for an analysis of the significant directional effects which were excluded). GNT was applied to all `r n_empiric_ns` effects. In contrast, QUID and OANOVA were used on subsets of `r n_empiric_ns_quid` and `r n_empiric_ns_oanova` of these effects, respectively (omitting five effects of signal detection sensitivity, d’, from both tests, and one additional RT interaction from the QUID analysis, as its current implementation only supports simple RT effects).
All tests agreed on finding no reliable evidence for non-directional unconscious effects. According to GNT, the prevalence statistic was zero in `r round(100*p_gnt_stat_zero,2)`% of the effects (maximal observed prevalence = `r ns_gnt_res %>% filter(is_sim == FALSE) %>% mutate(prev=round(100*gnt.stat, 2)) %>% pull(prev) %>% max()`%; see Fig. 2A), and the 95% one-sided CI included $\alpha=5$% in all of them. Hence, for all effects the prevalence of effects did not exceed the expected rate under the global null hypothesis. Similarly, for both QUID and the OANOVA tests, no single $BF$ or p-value revealed evidence for an effect (maximal $BF_{10}$ = `r round(ns_quid_res %>% filter(is_sim == FALSE) %>% pull(quid_bf) %>% max(), 2)` and all p-values > `r oanova_alpha`; see Figure 2B, C). Notably, QUID obtained moderate evidence for the *global null* model in `r round(100*p_null_QUID,2)`% of the cases (see Fig. 2B). The remaining effects were inconclusive. Hence, for the effects collected here, in the case of unconscious processing, the three tests revealed a highly similar pattern of results, consistent with a strong interpretation of previously reported null results as revealing the genuine absence of a causal effect of unconsciously perceived stimuli on behaviour.

(ref:figure2caption) The results of applying the GNT (A), QUID (B), and OANOVA (C) tests to effects that produced null results in a non-parametric directional test and to simulated data (the Non-directional effect (ND) and Global Null (GN) simulations described above, presented as square-shaped markers). Effect labels^[Effect labels abbreviations (sorted alphabetically): BH = @benthien2021a, BM = @biderman2018a, C = @chien2022a, F = @faivre2014a, H = @hurme2020a, S = @skora2021a, SV = @stein2021a, Z = @zerweck2021a. For all labels, numbers denote effect indices within each study (see Supplementary Table 1 for the full mapping between labels and effects).] appear on the x-axis. Panel A: the estimated prevalence of an unconscious effect in each of the cases, using GNT [@donhauser2018a]. Segments depict the one-sided 95% CI ($CI_{95}$) for the prevalence estimate. The solid orange line indicates the expected prevalence of 5% of significant individual-level effects, given that individual effects were tested using $\alpha = .05$. Panel B: Bayes factors for the comparison between a random effects model that takes into account potential differences in effect signs and the global null model. White markers depict cases where moderate evidence for the global null model was found, while grey markers indicate inconclusive results. The dashed black line indicates a BF of 1 (no preference for either model), and the solid orange lines indicate a BF cutoff of `r bf_criteria`. Panel C: p-values obtained by the OANOVA test [@miller2018a]. Blue and grey markers indicate significant and non-significant results, respectively. For illustration purposes, BF and significance values are presented on a logarithmic scale on the y-axis.
```{r figure2, out.width="100%", fig.cap = "(ref:figure2caption)"}
knitr::include_graphics("figures//plots//ns_available_methods_res.png")
```

@@

Second, both the model comparison approach used in QUID and the OANOVA test necessarily assume a parametric model of the data, making specific assumptions of normality and equal within-individual variance. In simulations, we find that violations of this second assumption can have dramatic effects on the specificity and sensitivity of both tests (see Appendix B). This can be addressed by more complex models that are capable of handling different distribution families, but as model complexity grows, unwanted effects of assumption violations may become harder to spot and quantify. Hence, taking a non-parametric approach provides safer inferences when the form of the data-generating process is not fully known.

Lastly, since GNT is focused on the prevalence of effects, it begins with testing the significance of effects at the single subject level, thereby dichotomizing a continuous test statistic into one bit of information: significant or not. This dichotomization results in information loss and introduces an additional free parameter — the individual alpha level. This step is well justified when estimating population prevalence, but it is unnecessary for our purpose of detecting a non-directional effect at the population level. As we describe below, using a continuous participant-level statistic makes our test more sensitive (see Appendix C for a direct comparison between the two
approaches).

In the next section, we introduce a novel non-directional test that takes into account population heterogeneity to infer group-level effects. The test is both frequentist and non-parametric, which addresses the above issues. Similarly to the OANOVA test, it promises a control for long term error-rates, but unlike it, our test does not assume a parametric model of the data-generating process. Using a continuous within-participant summary statistic, it is also more statistically powerful than approaches that focus on a dichotomous notion of effect prevalence (see Appendix C).

@@

Our test assumes that a within-participant effect is convincing if it is consistently evident across different trials. To test this, we can split the trials of an individual into two random halves, and ask whether both halves show the same qualitative effect (e.g., for both halves, the performance in the congruent condition is higher than in the incongruent condition; see Fig. 3A). By doing this many times, we can measure how often the two halves agree. Following this strategy, we estimate the consistency of effect signs within each individual by measuring the frequency of consistent results across splits. Then, we compare the group-mean consistency score against a null distribution: 10,000 samples of group-level consistency scores, obtained after randomly shuffling the experimental condition labels within participants (here, to speed up the computational process, for each participant, 100 permutations were created, from which we randomly sampled a single permutation in each null distribution sample; @stelzer2013a). Hence, our null distribution reflects the expected consistency of within-participant effects when the dependent variable of no single participant is sensitive to the experimental manipulation. An easy-to-use implementation of the sign consistency test is available as part of the signcon R package ([https://github.com/mufcItay/signcon](https://github.com/mufcItay/signcon); see Appendix D for extensions of the test to use cases that diverge from simple mean difference between conditions).

(ref:figure3caption) A frequentist, non-parametric, test for sign consistency. Panel A: a schematic illustration, using the same conventions as in Figure 1 (C = congruent, I = incongruent).Participant-wise sign consistency is quantified as the proportion of random splits of experimental trials, for which both halves display the same qualitative effect (C>I or I>C). Group-level sign consistency is compared against a non-parametric null distribution to obtain a significance value. The left panel illustrates a subset of random splits from a hypothetical participant. The upper row illustrates the overall RT data for that participant, and each row below shows one split of the data, where for each half we compare the mean of congruent and incongruent RT distributions, to test if the direction of the difference in the two halves is consistent or not. Then, to determine if the group shows evidence for non-directional effects, the averaged consistency score *across participants* (plotted in green), which is the proportion of consistent splits across all splits, is compared to the null distribution (right panel). In this hypothetical case, the group does not show an effect, as the average score is well within the null distribution. Panel B: the results of applying the sign consistency test to effects that produced null results in a non-parametric directional test (N = `r n_empiric_ns`). Significant results, for which the estimated mean sign consistency score is greater than 95% of the null distribution, are marked in blue. As in Figure 2, the x-axis lists effect labels.
```{r figure3, out.width="100%", fig.cap = "(ref:figure3caption)"}
knitr::include_graphics("figures//figure3.png")
```

We quantified the average within-participant sign consistency of effects for which a directional test did not produce significant results (the same effects reported in Fig. 2). For each individual, sign consistency was defined as the percentage of consistent signs across 500 random splits. Effect scores were calculated using a predefined summary function (i.e., taking the average RT or calculating signal detection sensitivity, d’, in each condition, depending on the effect of interest). The results revealed a similar picture to the one provided by the previous analyses (see Figure 3B). First, for the simulated datasets, the sign consistency test obtained non-significant results in the *global null* scenario (M = `r round(100*sc_gn_stat,0)`%, p = `r apa_p(sc_gn_p)`), and detected an effect in *non-directional differences* scenario (M = `r round(100*sc_nd_stat,0)`%, p `r apa_p(sc_nd_p)`). Second, for the empirical datasets, the vast majority of cases did not show significant sign consistency, with two exceptions: first, an effect of an unconsciously presented cue on wagering decisions (@skora2021a; M = `r round(100* skora_1_res %>% pull(signcon.statistic) %>% first(),0)`%, p = `r apa_p(skora_1_res %>% pull(signcon.p) %>% first())`), and second, a scene-object congruency effect (@biderman2018a; M = `r round(100* bm_3_res %>% pull(signcon.statistic) %>% first(),0)`%, p = `r apa_p(bm_3_res %>% pull(signcon.p) %>% first())`). Although the two effects were not detected by GNT, the prevalence of observed proportion of individual-level effects was above zero for both (`r round(100*skora_1_res %>% pull(gnt.stat) %>% first(),2)`% and `r round(100*bm_3_res %>% pull(gnt.stat) %>% first(),2)`%, respectively). Thus, despite some evidence for sign consistency, the overall picture remained the same, hinting at minimal qualitative inter-individual differences in unconscious processing.

```{r sc_power_tests, include=FALSE}
individual_effect_alpha <- .05
get_expected_prevalence <- function(prevalence, power, alpha) {
  return(prevalence * power + (1-prevalence) * alpha)
}
full_prevalence_exp_rate <- get_expected_prevalence(1, apndx_F_median_cnd_power, individual_effect_alpha) 
half_prevalence_exp_rate <- get_expected_prevalence(.5, apndx_F_median_cnd_power, individual_effect_alpha)

test_full_prevalence <- binom.test(sc_sig_n, n_empiric_ns, full_prevalence_exp_rate, alternative = 'less')
test_half_prevalence <- binom.test(sc_sig_n, n_empiric_ns, half_prevalence_exp_rate, alternative = 'less')
```
Together, four different analysis methods support the conclusion that by and large, unconscious priming effects are not masked by individual differences. Yet one can still claim that these statistical tests are simply not sensitive enough to detect qualitatively variable, non-directional effects, even when those exist. To test this claim, we conducted two additional analyses: First, we used simulations to estimate the sensitivity of our solution to non-directional effects with various effect sizes, determined according to previous analyses on unconscious processing [@meyen2022a] and cognitive control [@rouder2023a]. The results corroborated the concerns for lack of power when using common unconscious processing settings of the number of participants and trials (see @baker2021 for a detailed analysis of the contribution of both factors to power). Yet, we conducted further analysis showing that given the (low) power estimate we found, the number of significant effects obtained by the sign consistency test would be surprisingly low if an effect existed in all or even half of the datasets (see Appendix F). Second, to provide positive-control for these methods and show that they can be used to reveal such hidden effects in other fields, we collected additional, openly accessible, datasets from studies conducted in different fields of research within experimental psychology. We then used our non-parametric test on these datasets, demonstrating its potential benefit in determining whether a null result at the group level hides true, but variable, effects at the individual participant level.

@@

We used the sign consistency test to expose hidden effects that were not revealed by standard directional tests in various fields of research (see Appendix E for the same analysis using the three other tests). To that end, we exhausted all data from different open-access databases (the Confidence Database [@rahnev2020a], the Reproducibility Project [@collaboration2015a], and the Classic Visual Search Effects open dataset [@adam2021a]). We also used social media to ask for previously reported null effects, using the same inclusion criteria from the unconscious processing studies analysis, detailed above. Again, effects that were significant according to a non-parametric, directional sign-flipping test on the population mean were filtered out. Overall, we collected data associated with `r N_ns_no_UC` non-significant effects (`r N_ns_Confidence` from the Confidence Database, `r apa_num(as.integer(N_ns_Reproducibility), numerals = FALSE)` from the Reproducibility Project, `r apa_num(as.integer(N_ns_Attention), numerals = FALSE)` from the Classic Visual Search Effects open dataset and `r apa_num(as.integer(N_ns_Social),numerals = FALSE)` from the social media query). In all cases, participants were excluded for having fewer than five trials per experimental condition and/or zero variance in the dependent variable.

@@

To facilitate the adoption of this non-directional approach in experimental psychology, we release with this paper an R package with a simple-to-use implementation of our error-controlled and non-parametric sign consistency test ([https://github.com/mufcItay/signcon](https://github.com/mufcItay/signcon)). We note that unlike directional tests, the validity of the sign consistency test (and more generally, non-directional tests) depends on counterbalancing of confounding variables not only across participants, but also across trials within a single participant. We recommend using this test to complement standard, directional tests, taking into account the effect of additional tests on the family-wise error rate. Furthermore, although the test revealed effects in various domains, special attention should be given to statistical power when collecting data for a non-directional test, considering both the number of participants and the number of trials per participant. This is especially important when the effect size of interest is small, as is clearly the case in unconscious processing studies (more generally, when the relation between true variability between participants and measurement error is small; see @rouder2023a and Appendix F). Given proper use, the test should be particularly useful in interpreting null findings at the group level (see Box A for a more detailed description of best-practice recommendations for non-directional testing). This seems highly relevant to the field of unconscious processing, where null results are becoming more prevalent, and carry theoretical significance as hinting at possible functional roles for conscious processing.

@@

## Data and code availability
  All data, and analysis code are available at [https://github.com/mufcItay/NDT](https://github.com/mufcItay/NDT), using `r cite_r(file = "r-references.bib")`.

@@

\newpage
# Supplementary Table 1 {#STable-1}
``` {r supptable-1}
# create the supplementary table by code
Study <- c('Biderman & Mudrik, 2018', 'Faivre et al., 2014', 'Stein & Peelen, 2021', 
'Zerweck et al., 2021', 'Benthien & Hesselmann, 2021', 'Hurme et al., 2020', 
'Skora et al., 2021', 'Chien et al., 2022')
Labels <- c('BM1-3', 'F1-8', 'SVP1-5','Z1-7','BH1','H1-4','S1-2','C1-3')
Topic <- c('Scene congruency', 'Multisensory integration', 'Location effects + PAS (detection)',
'Numerical Priming', 'Numerical Priming', 'Colours', 'Instrumental Learning', 'Semantic priming')
Paradigm <- c('Masking', 'Masking', 'CFS', 'Masking', 'CFS', 'TMS + Metacontrast Masking', 'Masking', 'CFS')
DV <- c('RT', 'RT', 'd\'', 'RT', 'RT', 'RT', 'd\'', 'RT')
Notes <- c('Replication study. For all experiments, log(RT) was used in the original analysis',
'Four experiments, with two effects in each experiment (identical/different targets). For all experiments, log(RT) was used in the original analysis',
'Two experiments (3&4 in the paper), measuring effects in different prime-mask SOAs',
'Two experiments (2 and 3 in the original paper), measuring effects in different SOA / Contrast conditions',
'Interaction effect - prime congruency X location certainty',
'Redundant target effect (TMS / Masking X Blue / Red)',
'Regression to the mean as a confound according to authors',
'Word, Picture, and trait discrimination tasks')
studies_df <- data.frame(Study, Labels, Topic, Paradigm, DV, Notes)
# print the table according to apa style using a small font size
apa_table(
  studies_df,
  caption = "Unconscious processing effects metadata",
  font_size = 'footnotesize')
```

# Supplementary Figure 1 - GNT, QUID and OANOVA results for datasets showing a directional effect {#SFig-1}

(ref:sfigure1caption) Supplementary Figure 1. The results of applying the GNT (A), QUID (B) and OANOVA (C) tests to effects that produced significant results in a non-parametric directional test. Same conventions as Fig. 2.
```{r Sfigure1, out.width="100%", fig.cap="(ref:sfigure1caption)", warning = FALSE}
knitr::include_graphics("figures//plots//effect_available_methods_res.png")
```

# Supplementary Figure 2 - Sign consistency test results for datasets showing a directional effect
(ref:sfigure2caption) Supplementary Figure 2. The results of applying the sign consistency test to significant directional effects (N = 7). The x-axis lists effect labels. Same conventions as Fig. 3.
```{r sfigure2, out.width="100%", fig.cap = "(ref:sfigure2caption)", warning = FALSE}
knitr::include_graphics("figures//plots//effect_plt_sign_con__res.png")
```

@@

# Appendix D. Extending the sign consistency test to additional use cases {#Apndx-D}

We provide two code examples to demonstrate how to extend the sign-consistency test implemented in the signcon R package ([https://github.com/mufcItay/signcon](https://github.com/mufcItay/signcon)) to common use case: D.1. A 2 X 2 interaction, and D.2. calculating SDT's d'. For both examples we will use simulated trial-level data, mimicking a 2X2 within-participant design.

>>>>

authornote: |
  All data except for the datasets from Machado et al., 2007, 2009, for which participants did not consent to having their data shared online, and all analysis code are available at https://github.com/mufcItay/NDT. This study's analysis was not pre-registered.

@@

# appendix B is a table with a description of all unconscious processing studies, and it is created below in code

# appendix C is created by running source('figures\\fig2_3.R')

# the analysis and simulations for appendix D.
# to run the analysis again, uncomment the (commented) run_appendixD line and comment the read.csv line which reads the results of an existing simulations run.
source('appendix\\appendix_D.R')
# appendixD_res <- run_appendixD(apndxD_conf_FAs, apndxD_conf_sensitivity)
appendixD_res <- read.csv('appendix\\results\\2024-01-07_23_39_16Appendix_D_QUID_OANOVA.csv')

# the analysis and simulations for appendix E.
# to run the analysis again, uncomment the (commented) run_appendixE line and comment the read.csv line which reads the results of an existing simulations run (keeping the save_plot_appendixE line.)
source('appendix\\appendix_E.R')
# appendixE_res <- run_appendixE(apndx_e_conf)
appendixE_res <- read.csv('appendix\\results\\2023-12-29_18_48_11Appendix_E.csv')
save_plot_appendixE(appendixE_res)

# appendix F was not implemented in an external script.
# see the implementation below.

# appendix G is created by running source('figures\\fig2_3.R')

# the analysis and simulations for appendix H.
# to run the analysis again, uncomment the (commented) run_appendixH line and comment the read.csv line which reads the results of an existing simulations run.
source('appendix\\appendix_H.R')
apndxH_N_p <- N_p
apndxH_N_t <- N_t
# appendixH_res <- run_appendixH(apndx_e_conf)
appendixH_res <- read.csv('appendix\\results\\2024-01-09_16_46_09Appendix_H.csv')
apndx_H_median_cnd_power <- appendixH_res %>%
  filter(N_p == median(appendixH_res$N_p), 
         N_t == median(appendixH_res$N_t),
         sds_ratio == median(appendixH_res$sds_ratio)) %>%
  pull(Power)

# the materials for appendix I are generated by sourcing the
# 'figure4.R' script, and the its figure was created with minor
# edits applied to the output files

@@

We had three inclusion criteria: first, since all probed tests require trial-level data, only open-access datasets providing such data were included. Second, the independent variable of all studies had to be manipulated within single participants. And third, at least one non-significant effect was reported in the original study. Overall, this search strategy yielded data associated with `r n_empiric_ns` null effects (see Appendix B for details about all effects), 21 focusing on differences in RT and 5 on differences in signal detection sensitivity, d’ [@green1966a]. We used the criteria set by the original authors for demonstrating unawareness (e.g., using objective and/or subjective measures of awareness), and a two-sided non-parametric sign-flipping test on the population mean for filtering out significant priming effects^[Across all RT effects, our analysis used raw RT scores, and thus our results diverged from the original results when log transformations were used (see the notes column in Appendix B for details)]. Finally, we excluded participants with fewer than five trials per experimental condition and/or zero variance in the dependent variable (e.g., when accuracy was measured). Together, these data allowed us to reexamine null unconscious processing effects using a non-directional approach that takes into account the potential for differences in effect signs when testing for group-level effects. We accordingly asked whether true effects of unconscious processing were masked by population heterogeneity in effect signs.

To that end, the effects of interest were tested using GNT, QUID, and the OANOVA tests (see Appendix C for an analysis of the significant directional effects which were excluded). GNT was applied to all `r n_empiric_ns` effects. In contrast, QUID and OANOVA were used on subsets of `r n_empiric_ns_quid` and `r n_empiric_ns_oanova` of these effects, respectively (omitting five effects of signal detection sensitivity, d’, from both tests, and one additional RT interaction from the QUID analysis, as its current implementation only supports simple RT effects).
All tests agreed on finding no reliable evidence for non-directional unconscious effects. According to GNT, the prevalence statistic was zero in `r round(100*p_gnt_stat_zero,2)`% of the effects (maximal observed prevalence = `r ns_gnt_res %>% filter(is_sim == FALSE) %>% mutate(prev=round(100*gnt.stat, 2)) %>% pull(prev) %>% max()`%; see Fig. 2A), and the 95% one-sided CI included $\alpha=5$% in all of them. Hence, for all effects the prevalence of effects did not exceed the expected rate under the global null hypothesis. Similarly, for both QUID and the OANOVA tests, no single $BF$ or p-value revealed evidence for an effect (maximal $BF_{10}$ = `r round(ns_quid_res %>% filter(is_sim == FALSE) %>% pull(quid_bf) %>% max(), 2)` and all p-values > `r oanova_alpha`; see Figure 2B, C). Notably, QUID obtained moderate evidence for the *global null* model in `r round(100*p_null_QUID,2)`% of the cases (see Fig. 2B). The remaining effects were inconclusive. Hence, for the effects collected here, in the case of unconscious processing, the three tests revealed a highly similar pattern of results, consistent with a strong interpretation of previously reported null results as revealing the genuine absence of a causal effect of unconsciously perceived stimuli on behaviour.

(ref:figure2caption) The results of applying the GNT (A), QUID (B), and OANOVA (C) tests to effects that produced null results in a non-parametric directional test and to simulated data (the Non-directional effect (ND) and Global Null (GN) simulations described above, presented as square-shaped markers). Effect labels^[Effect labels abbreviations (sorted alphabetically): BH = @benthien2021a, BM = @biderman2018a, C = @chien2022a, F = @faivre2014a, H = @hurme2020a, S = @skora2021a, SV = @stein2021a, Z = @zerweck2021a. For all labels, numbers denote effect indices within each study (see Appendix B for the full mapping between labels and effects).] appear on the x-axis. Panel A: the estimated prevalence of an unconscious effect in each of the cases, using GNT [@donhauser2018a]. Segments depict the one-sided 95% CI ($CI_{95}$) for the prevalence estimate. The solid orange line indicates the expected prevalence of 5% of significant individual-level effects, given that individual effects were tested using $\alpha = .05$. Panel B: Bayes factors for the comparison between a random effects model that takes into account potential differences in effect signs and the global null model. White markers depict cases where moderate evidence for the global null model was found, while grey markers indicate inconclusive results. The dashed black line indicates a BF of 1 (no preference for either model), and the solid orange lines indicate a BF cutoff of `r bf_criteria`. Panel C: p-values obtained by the OANOVA test [@miller2018a]. Blue and grey markers indicate significant and non-significant results, respectively. For illustration purposes, BF and significance values are presented on a logarithmic scale on the y-axis.
```{r figure2, out.width="100%", fig.cap = "(ref:figure2caption)"}
knitr::include_graphics("figures//plots//ns_available_methods_res.png")
```

@@

Second, both the model comparison approach used in QUID and the OANOVA test necessarily assume a parametric model of the data, making specific assumptions of normality and equal within-individual variance. In simulations, we find that violations of this second assumption can have dramatic effects on the specificity and sensitivity of both tests (see Appendix D). This can be addressed by more complex models that are capable of handling different distribution families, but as model complexity grows, unwanted effects of assumption violations may become harder to spot and quantify. Hence, taking a non-parametric approach provides safer inferences when the form of the data-generating process is not fully known.

Lastly, since GNT is focused on the prevalence of effects, it begins with testing the significance of effects at the single subject level, thereby dichotomizing a continuous test statistic into one bit of information: significant or not. This dichotomization results in information loss and introduces an additional free parameter — the individual alpha level. This step is well justified when estimating population prevalence, but it is unnecessary for our purpose of detecting a non-directional effect at the population level. As we describe below, using a continuous participant-level statistic makes our test more sensitive (see Appendix E for a direct comparison between the two
approaches).

In the next section, we introduce a novel non-directional test that takes into account population heterogeneity to infer group-level effects. The test is both frequentist and non-parametric, which addresses the above issues. Similarly to the OANOVA test, it promises a control for long term error-rates, but unlike it, our test does not assume a parametric model of the data-generating process. Using a continuous within-participant summary statistic, it is also more statistically powerful than approaches that focus on a dichotomous notion of effect prevalence (see Appendix E).

@@

Our test assumes that a within-participant effect is convincing if it is consistently evident across different trials. To test this, we can split the trials of an individual into two random halves, and ask whether both halves show the same qualitative effect (e.g., for both halves, the performance in the congruent condition is higher than in the incongruent condition; see Fig. 3A). By doing this many times, we can measure how often the two halves agree. Following this strategy, we estimate the consistency of effect signs within each individual by measuring the frequency of consistent results across splits. Then, we compare the group-mean consistency score against a null distribution: 10,000 samples of group-level consistency scores, obtained after randomly shuffling the experimental condition labels within participants (here, to speed up the computational process, for each participant, 100 permutations were created, from which we randomly sampled a single permutation in each null distribution sample; @stelzer2013a). Hence, our null distribution reflects the expected consistency of within-participant effects when the dependent variable of no single participant is sensitive to the experimental manipulation. An easy-to-use implementation of the sign consistency test is available as part of the signcon R package ([https://github.com/mufcItay/signcon](https://github.com/mufcItay/signcon); see Appendix F for extensions of the test to use cases that diverge from simple mean difference between conditions).

(ref:figure3caption) A frequentist, non-parametric, test for sign consistency. Panel A: a schematic illustration, using the same conventions as in Figure 1 (C = congruent, I = incongruent). Participant-wise sign consistency is quantified as the proportion of random splits of experimental trials, for which both halves display the same qualitative effect (C>I or I>C). Group-level sign consistency is compared against a non-parametric null distribution to obtain a significance value. The left panel illustrates a subset of random splits from a hypothetical participant. The upper row illustrates the overall RT data for that participant, and each row below shows one split of the data, where for each half we compare the mean of congruent and incongruent RT distributions, to test if the direction of the difference in the two halves is consistent or not. Then, to determine if the group shows evidence for non-directional effects, the averaged consistency score *across participants* (plotted in green), which is the proportion of consistent splits across all splits, is compared to the null distribution (right panel). In this hypothetical case, the group does not show an effect, as the average score is well within the null distribution. Panel B: the results of applying the sign consistency test to effects that produced null results in a non-parametric directional test (N = `r n_empiric_ns`). Significant results, for which the estimated mean sign consistency score is greater than 95% of the null distribution, are marked in blue. As in Figure 2, the x-axis lists effect labels.
```{r figure3, out.width="100%", fig.cap = "(ref:figure3caption)"}
knitr::include_graphics("figures//figure3.png")
```

We quantified the average within-participant sign consistency of effects for which a directional test did not produce significant results (the same effects reported in Fig. 2; see Appendix G for an analysis of the excluded significant directional effects). For each individual, sign consistency was defined as the percentage of consistent signs across 500 random splits. Effect scores were calculated using a predefined summary function (i.e., taking the average RT or calculating signal detection sensitivity, d’, in each condition, depending on the effect of interest). The results revealed a similar picture to the one provided by the previous analyses (see Figure 3B). First, for the simulated datasets, the sign consistency test obtained non-significant results in the *global null* scenario (M = `r round(100*sc_gn_stat,0)`%, p = `r apa_p(sc_gn_p)`), and detected an effect in *non-directional differences* scenario (M = `r round(100*sc_nd_stat,0)`%, p `r apa_p(sc_nd_p)`). Second, for the empirical datasets, the vast majority of cases did not show significant sign consistency, with two exceptions: first, an effect of an unconsciously presented cue on wagering decisions (@skora2021a; M = `r round(100* skora_1_res %>% pull(signcon.statistic) %>% first(),0)`%, p = `r apa_p(skora_1_res %>% pull(signcon.p) %>% first())`), and second, a scene-object congruency effect (@biderman2018a; M = `r round(100* bm_3_res %>% pull(signcon.statistic) %>% first(),0)`%, p = `r apa_p(bm_3_res %>% pull(signcon.p) %>% first())`). Although the two effects were not detected by GNT, the prevalence of observed proportion of individual-level effects was above zero for both (`r round(100*skora_1_res %>% pull(gnt.stat) %>% first(),2)`% and `r round(100*bm_3_res %>% pull(gnt.stat) %>% first(),2)`%, respectively). Thus, despite some evidence for sign consistency, the overall picture remained the same, hinting at minimal qualitative inter-individual differences in unconscious processing.

```{r sc_power_tests, include=FALSE}
individual_effect_alpha <- .05
get_expected_prevalence <- function(prevalence, power, alpha) {
  return(prevalence * power + (1-prevalence) * alpha)
}
full_prevalence_exp_rate <- get_expected_prevalence(1, apndx_H_median_cnd_power, individual_effect_alpha) 
half_prevalence_exp_rate <- get_expected_prevalence(.5, apndx_H_median_cnd_power, individual_effect_alpha)

test_full_prevalence <- binom.test(sc_sig_n, n_empiric_ns, full_prevalence_exp_rate, alternative = 'less')
test_half_prevalence <- binom.test(sc_sig_n, n_empiric_ns, half_prevalence_exp_rate, alternative = 'less')
```
Together, four different analysis methods support the conclusion that by and large, unconscious priming effects are not masked by individual differences. Yet one can still claim that these statistical tests are simply not sensitive enough to detect qualitatively variable, non-directional effects, even when those exist. To test this claim, we conducted two additional analyses: First, we used simulations to estimate the sensitivity of our solution to non-directional effects with various effect sizes, determined according to previous analyses on unconscious processing [@meyen2022a] and cognitive control [@rouder2023a]. The results corroborated the concerns for lack of power when using common unconscious processing settings of the number of participants and trials (see @baker2021 for a detailed analysis of the contribution of both factors to power). Yet, we conducted further analysis showing that given the (low) power estimate we found, the number of significant effects obtained by the sign consistency test would be surprisingly low if an effect existed in all or even half of the datasets (see Appendix H). Second, to provide positive-control for these methods and show that they can be used to reveal such hidden effects in other fields, we collected additional, openly accessible, datasets from studies conducted in different fields of research within experimental psychology. We then used our non-parametric test on these datasets, demonstrating its potential benefit in determining whether a null result at the group level hides true, but variable, effects at the individual participant level.

@@

We used the sign consistency test to expose hidden effects that were not revealed by standard directional tests in various fields of research (see Appendix I for the same analysis using the three other tests). To that end, we exhausted all data from different open-access databases (the Confidence Database [@rahnev2020a], the Reproducibility Project [@collaboration2015a], and the Classic Visual Search Effects open dataset [@adam2021a]). We also used social media to ask for previously reported null effects, using the same inclusion criteria from the unconscious processing studies analysis, detailed above. Again, effects that were significant according to a non-parametric, directional sign-flipping test on the population mean were filtered out. Overall, we collected data associated with `r N_ns_no_UC` non-significant effects (`r N_ns_Confidence` from the Confidence Database, `r apa_num(as.integer(N_ns_Reproducibility), numerals = FALSE)` from the Reproducibility Project, `r apa_num(as.integer(N_ns_Attention), numerals = FALSE)` from the Classic Visual Search Effects open dataset and `r apa_num(as.integer(N_ns_Social),numerals = FALSE)` from the social media query). In all cases, participants were excluded for having fewer than five trials per experimental condition and/or zero variance in the dependent variable.

@@

To facilitate the adoption of this non-directional approach in experimental psychology, we release with this paper an R package with a simple-to-use implementation of our error-controlled and non-parametric sign consistency test ([https://github.com/mufcItay/signcon](https://github.com/mufcItay/signcon)). We note that unlike directional tests, the validity of the sign consistency test (and more generally, non-directional tests) depends on counterbalancing of confounding variables not only across participants, but also across trials within a single participant. We recommend using this test to complement standard, directional tests, taking into account the effect of additional tests on the family-wise error rate. Furthermore, although the test revealed effects in various domains, special attention should be given to statistical power when collecting data for a non-directional test, considering both the number of participants and the number of trials per participant. This is especially important when the effect size of interest is small, as is clearly the case in unconscious processing studies (more generally, when the relation between true variability between participants and measurement error is small; see @rouder2023a and Appendix H). Given proper use, the test should be particularly useful in interpreting null findings at the group level (see Box A for a more detailed description of best-practice recommendations for non-directional testing). This seems highly relevant to the field of unconscious processing, where null results are becoming more prevalent, and carry theoretical significance as hinting at possible functional roles for conscious processing.

@@

## Data and code availability
  All data except for the datasets from Machado et al., 2007, 2009, for which participants did not consent to having their data shared online, and all analysis code are available at [https://github.com/mufcItay/NDT](https://github.com/mufcItay/NDT), using `r cite_r(file = "r-references.bib")`.

@@

@@

# Appendix F. Extending the sign consistency test to additional use cases {#Apndx-F}

We provide two code examples to demonstrate how to extend the sign-consistency test implemented in the signcon R package ([https://github.com/mufcItay/signcon](https://github.com/mufcItay/signcon)) to common use case: F.1. A 2 X 2 interaction, and F.2. calculating SDT's d'. For both examples we will use simulated trial-level data, mimicking a 2X2 within-participant design.
